NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Molecular Dynamics Simulations

### 2023-04-13 MANUSCRIPT/manuscript.tex

Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} which are unevenly distributed in the membrane plane and across its leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}. A key player driving the lateral heterogeneity is cholesterol (CHOL), which is present at concentrations between $\sim$10~mol-\% (endoplasmic reticulum) and up to $\sim$50~mol-\% (plasma membrane) \cite{van2008membrane}. CHOL has the unique ability to order neighbouring lipids and thus induce the liquid ordered (L\textsubscript{o}) phase in model membranes \cite{mouritsen2004s,ipsen87,kinnunen91,rog2009ordering}. In the cellular setting, the interaction between phospholipids and CHOL is associated with the formation of lipid rafts and nanodomains \cite{Simons97,cebecauer2018membrane}. This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation} \textit{via} the hydrophobic mismatch mechanism \cite{killian1998hydrophobic}, in addition to the direct modulation of protein function \cite{gimpl2016interaction,guixa2017membrane}.

While detailed experimental information of the interactions among phospholipids and CHOL is relatively sparse, atomistic resolution molecular dynamics (MD) simulations have been widely applied to provide a detailed view of the lateral organization of lipid bilayers as well as lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed}. However, the trustworthiness of any prediction from an MD simulation relies on the quality of the used model. Fortunately, in addition to constant refinement of existing parameters, atomistic \emph{force fields} have also seen the expansion to other classes of biomolecules, including lipids, sugars, and nucleic acids \cite{leonard2019developing,marrink2019computational}. The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have growing libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the model by Maciejewski and Rog (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig15b}, respectively. Moreover, OPLS3e \cite{harder2016opls3} contains parameters for lipids, but CHOL is currently not included.

Thanks to the compatibility with protein force fields, the lipid models can be used in studies of membrane proteins or protein--membrane interactions in addition to sole lipid membranes. Notably, simulations using CHARMM36, Lipid17, and Slipids are now readily set up using CHARMM-GUI for multiple simulation engines \cite{lee16,lee2020charmm}. Despite this ease-of-use, the choice of a force field for an MD simulation can be challenging, as each model comes with its own set of pros and cons. Unfortunately, careful comparisons of force field quality, especially for capturing the features of lipid mixtures, are scarce in the literature. This largely results from the lack of availability of systematic, comprehensive, and quantitative datasets from both experiments and simulations.

We have previously performed such comparisons for the structure of phosphatidylcholine (PC) lipids \cite{botan15}, for PC--ion interactions \cite{catte2016molecular}, for phosphatidylserine (PS) lipids \cite{antila2019headgroup}, and for phoshatidylethanolamine (PE) and phosphatidylglycerol (PG) lipids \cite{bacle2021inverse}. Here, we extend this work to mixtures of PC and CHOL. We compare the ability of popular force fields to reproduce effects of CHOL on the structure and dynamics of a PC lipid bilayer, and thereby set guidelines for future efforts to validate intermolecular interactions in binary and more complex systems. Moreover, if some of the studied force fields satisfactorily reproduce experimental behavior at least in some range of CHOL concentrations, the molecular level picture provided by it can be assumed to mimic reality in this concentration regime.

Our earlier work has demonstrated that the structural ensembles of lipids from an MD simulation can be directly compared to those from solid state NMR experiments \cite{botan15,catte2016molecular,antila2019headgroup,bacle2021inverse}. This approach based on order parameters is quantitative, has small error estimates, is highly reproducible, and does not suffer from perturbation of probes used in many other techniques. Thus, we compare our simulation results with different force fields to the systematic NMR data set of PC C--H bond order parameters with varying CHOL concentrations \cite{ferreira13}. In addition, we measured form factors for a range of PC--CHOL mixtures using small angle X-ray scattering, and used them to subsequently model the electron density profiles across the lipid bilayer. The form factor and electron density profiles can be compared with the same profiles calculated from the simulation data. Whereas NMR provides data on the conformations of individual lipids, form factors from scattering captures the details of the structure of the entire lipid bilayer \cite{ollila16}.

One limitation in comparing simulations of lipid mixtures with experimental data is that the various experiments are performed at different concentrations, and thus mimicking all compositions in simulations is tedious. Here, we demonstrate that if the simulations and experiments sample multiple CHOL concentrations, the order parameters and electron density profiles can be readily interpolated to intermediate CHOL concentrations. This enables the comparison of data sets that are not extracted at the exact same CHOL concentrations, as is usually the case, especially if multiple techniques are combined in the comparison.

We also evaluated whether the simulation force fields reproduce the dependence of lateral diffusion coefficients on cholesterol measured using pulsed field gradient (PFG) NMR \cite{filippov2003effect,filippov2003influence}. This comparison is enabled by the recent theoretical work that allows for the elimination of finite-size effects in MD simulations and thus a quantitative comparison with experiment \cite{vogele2016divergent,vogele2018hydrodynamics}. With the structural and dynamic comparisons established, we estimated the quality of each force field at different CHOL concentrations. We then analyzed the molecular organization in the best-performing force field at low and high CHOL range, as we expect them to best reflect this organization in reality.

@@

\subsection{X-ray Scattering Experiments}

We measured small angle X-ray scattering (SAXS) data on multilamellar vesicles (MLVs) composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments. The data were obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons at $T=300$~K and analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}. All lipids were purchased from Avanti Polar Lipids  (Alabaster, AL, USA) as dry powders and were used without any further purification. All other chemicals were obtained in pro analysis quality from Lactan (Graz, Austria). Fully hydrated MLVs were prepared using rapid solvent exchange as detailed  previously \cite{rieder2015optimizing,belivcka2017high}. This avoids the precipitation of cholesterol crystallites at high concentration \cite{buboltz1999novel} yielding non-phase separated samples up to 50~mol-\% cholesterol content.

The data from MLVs contain the structure factor (the crystalline lattice) and form factor in a convoluted fashion, yet by fitting the scattered intensity data we obtain both contributions. In this work, only the form factors are discussed. The electron density profiles have been modelled in terms of the SDP model \cite{heberle12,Kucerka08a,kucerka12}, \textit{i.e.} volume distribution functions are modelled by individual Gaussians or error functions. CHOL is particularly challenging to describe within the SDP model, and therefore form factors are often compared between simulation and experiment \cite{Kucerka08b} instead of density profiles, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}. Here, cholesterol is accounted for by two Gaussians, as proposed by Jianjun Pan (private communication), and used in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}. 
%Additional figures show the volume distribution functions and the resulting electron density profiles.
%Authors to consult and potentially include in publications using this data: Peter Heftberger (peter.heftberger@gmx.at), Georg Pabst (georg.pabst@uni-graz.at)

@@

We performed MD simulations of a pure POPC membrane as well as five POPC/CHOL mixtures with CHOL content ranging from 11 to 47~mol-\%. In order to eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes. The small (64 POPC molecules) and medium (256 POPC molecules) systems were only used in the diffusion analysis, whereas all other analyses were performed on the large membranes (1024 POPC molecules). The number of POPC molecules was kept constant across the different CHOL concentrations. All membranes were solvated by 50 waters per lipid (POPC or CHOL). The simulations were performed using four commonly used force fields, namely CHARMM36 \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig15b} models. All simulations were 1~\textmu{}s long, totaling 72~\textmu{}s, and performed using GROMACS version 2020 or 2021 \cite{pall2020heterogeneous}. The used simulation parameters are provided in Table~\ref{SItab:params}, and the simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749 (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog).

@@

\paragraph{Electron Density Profiles and Bilayer Dimensions:} The form factors and electron density profiles were calculated using the \texttt{form\_factor.py} script in the NMRlipids Databank (\url{https://github.com/NMRLipids/Databank/}) after assigning all atom types with valence electron counts. The thicknesses were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data. Area per phospholipid was obtained by dividing the area of the bilayer by the number of phospholipids in one leaflet.

\paragraph{Order Parameters:} The C--H order parameters were calculated as
%
\begin{align}
    S_\mathrm{CD}=\left\langle\frac{3\cos^2\theta-1}{2}\right\rangle,
\end{align}
%
where $\theta$ is the angle between the C--H vector and the $z$ axis normal to the bilayer plane. In the POPC acyl chains, the values for the 2 (3) hydrogens of CH$_2$ (CH$_3$) groups were averaged, since these groups rotate freely and thus the order parameters are identical for both (all) hydrogens in experiments. In simulations, the values are also very similar, justifying this approach. An exception to this are the hydrogens bound to the 2\textsuperscript{nd} carbon in the oleate chain; they lack rotational averaging in both both simulations and experiments, and were thus treated separately in our analyses. The C--H order parameters for the POPC head group were calculated similarly, but values for both hydrogens are shown separately for all CH$_2$ groups, \textit{i.e.} no averaging was performed.

@@

The diffusion coefficients extracted from the three simulation box sizes were fitted with
%
\begin{align}
	D_\mathrm{PBC}\approx D_\infty+\frac{k_\mathrm{B}T}{4\pi\mu_\mathrm{m}h}\frac{\ln\left[L/\left(L_\mathrm{SD}+1.565H\right)\right]-1.713}{1+H/L_\mathrm{SD}},
	\label{eq:sdlatpbc}
\end{align}
%
where $D_\infty$ is the lateral diffusion coefficient in an infinite system, $h$ is the hydrodynamic thickness of the membrane, $k_\mathrm{B}T$ is the Boltzmann constant, $T$ is the temperature, $mu_\mathrm{m}$ and $mu_\mathrm{f}$ are shear viscosities of the membrane and the fluid (water), respectively. $H$ is half the thickness of the water layer and $L_\mathrm{SD}=\frac{h\mu_\mathrm{m}}{2\mu_\mathrm{f}}$ the Saffmann--Delbr\"{u}ck length \cite{vogele2018hydrodynamics}. The inter-leaflet friction coefficient does not appear in Eq.~\eqref{eq:sdlatpbc} as we expect it to be infinite, which was found to be a valid assumption for lipid bilayers \cite{vogele2018hydrodynamics}. The water viscosity value of 0.3228~mPa$\cdot$s was interpolated to 298~K from the values for CHARMM TIP3P in Ref.~\citenum{ong2019temperature} and used for all simulations (CHARMM TIP3P and normal TIP3P differ by $\sim$2--3\%). The final simulation box dimensions were used for $L$ and $H$. For $h$ and $H$ ($=(L_z-h)/2$, the average values from the three systems sizes were used.

@@

\paragraph{Quantitative Quality Measure for Lipid Bilayer Structure:} The deviations (in \%) from experimental values across CHOL concentrations were used to quantify the quality of the lipid force fields. 
%The density profiles (as a function of distance from membrane core along the membrane normal), 
The 1\textsuperscript{st} and 2\textsuperscript{nd} form factor minima locations, the electron density profiles (as a function of normal distance from bilayer center), the oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains) and diffusion coefficients were interpolated to create continuous 2D matrices for the order parameters and density profiles, and a continuous 1D vector for the diffusion coefficients and form factor minima. For the 2D matrices, the absolute values of the differences of the matrices from simulations and experiments were calculated and averaged over the carbon atoms in the acyl chain (order parameters) or across the simulation system (density profiles). This resulted in 1D deviation vectors as a function of CHOL concentration. For diffusion coefficients and form factor minima, the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration. All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \%.

@@

\subsection{Cholesterol Condensing Effect is Manifested Differently in Different Force Fields}

Lipid bilayer dimensions can be experimentally accessed by measuring X-ray scattering form factor, which is related to the electron density along membrane normal \textit{via} a Fourier transform \cite{pan12,Heftberger15,Marquardt15,ollila16,??}. The form factor can be translated to an electron density profile, and area per lipid and bilayer thickness values can be extracted using the scattering density profile (SDP) model or its combination with MD simulations  \cite{Kucerka08a,pan12,Heftberger15,Marquardt15,??,doktorova2020molecular}. Sophisticated approaches for single component lipid bilayers are available but the data for multi-component systems are more difficult to interpret, and the used approaches involve severe approximations \cite{pan12,Heftberger15,Marquardt15,??}.

Thus, we compare here the form factors and the total electron density profiles which do not require the de-convolution of the electron density profiles into densities of the different molecular species or their segments. For reference, all scattering intensitities from experiment are shown in Figs.~\ref{SIfig:scattering1} and \ref{SIfig:scattering2}, the form factor profiles from experiment and simulation in Fig.~\ref{SIfig:scattering} and the density profiles from experiment and simulation in Fig.~\ref{SIfig:densprofs}.

As the absolute and relative lobe amplitudes depend on simulation box size (Fig.~\ref{SIfig:scatteringvssize}), the minima in the form factor profiles provide a robust comparison between simulation and experiment. Thus, we first traced the first two minima in the form factors, and they are plotted in Fig.~\ref{SIfig:ffminima}. As simulations and experiments were performed at somewhat different CHOL concentrations, we  interpolated the minima to the entire studied range of CHOL concentrations. The plots in Fig.~\ref{SIfig:ffminima} highlight that the addition of CHOL first shifts the first minima to smaller wave vector values in the experiment, which is reasonably well captured by the simulation force fields. CHARMM36 seems to be off more than the other three force fields. Above $\sim$25~mol-\% of CHOL, the location of the minimum shifts to larger wave vector values in the experiment, which is curiously not captured by any of the force fields. The experiment shows a steady shift of the second minimum to smaller wave vector values, and this is reproduced by all simulation force fields. Slipids and Lipid17 are generally in best agreement with the experiment, whereas MacRog and CHARMM36 show poorer agreement.

Moving on to the electron density profiles extracted from the form factors, we generated two-dimensional maps for the density profiles using interpolation of the density profiles in Fig.~\ref{SIfig:densprofs} (see Methods). These maps are shown in Fig.~\ref{fig:densmaps} for both simulations (top row) and experiment (bottom row, left column).

>>>>

Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} which are unevenly distributed in the membrane plane and across its leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}. A key player driving the lateral heterogeneity is cholesterol (CHOL), which is present at concentrations between $\sim$10~mol-\% (endoplasmic reticulum) and up to $\sim$50~mol-\% (plasma membrane) \cite{van2008membrane}. CHOL has the unique ability to order neighbouring lipids and thus induce the liquid ordered (L\textsubscript{o}) phase in model membranes \cite{mouritsen2004s,ipsen87,kinnunen91,rog2009ordering}. In the cellular setting, the interaction between phospholipids and CHOL is associated with the formation of lipid rafts and nanodomains \cite{Simons97,cebecauer2018membrane}. This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation}, %\textit{via} the hydrophobic mismatch mechanism \cite{killian1998hydrophobic}, 
in addition to the direct modulation of protein function \cite{gimpl2016interaction,guixa2017membrane}.

While the structure and dynamics of heterogeneous membranes are difficult to capture experimentally, atomistic resolution molecular dynamics (MD) simulations are widely used to provide a detailed view of complex membranes and the lateral organization of lipid bilayers driven by lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed,enkavi2019multiscale,marrink2019computational}. This is further facilitated by an increasing selection of available force fields with compatible lipid and protein parameters that enable simulations of complex membranes approaching realistic membrane mixtures.

The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have growing libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the model by Maciejewski and Rog (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig15b}, respectively. Notably, simulations using CHARMM36, Lipid17, and Slipids are now readily set up using CHARMM-GUI for multiple simulation engines that enables rapid set up of complex membrane simulations \cite{lee16,lee2020charmm}.

While simulating of complex membranes with cholesterol have become relatively straightforward task, estimating the trustworthiness of MD simulations for complex membranes remains a challenge.
%relies on the quality of the used model. 
%Despite this ease-of-use, the choice of a force field for an MD simulation can be challenging, as each model comes with its own set of pros and cons. Unfortunately, careful comparisons of force field quality, especially for capturing the features of lipid mixtures, are scarce in the literature. This largely results from the lack of availability of systematic, comprehensive, and quantitative datasets from both experiments and simulations.
%
Our earlier work has demonstrated that the conformational ensembles of lipids from an MD simulation can be evaluated against C-H bond order parameters from solid state NMR experiments \cite{botan15,ollila16,catte2016molecular,antila2019headgroup,bacle2021inverse}. 
%This approach based on order parameters is quantitative, has small error estimates, is highly reproducible, and does not suffer from perturbation of probes used in many other techniques. 
This approach has been useful for finding the best models for headgroups of phosphatidylcholine (PC)~\cite{botan15}, phosphatidylserine (PS)~\cite{antila2019headgroup}, phoshatidylethanolamine (PE) \cite{bacle2021inverse} and phosphatidylglycerol (PG) lipids \cite{bacle2021inverse}, for evaluating and improving membrane interactions with ions~\cite{catte2016molecular,antila2019headgroup,bacle2021inverse,melcr18,melcr19} and small molecules~\cite{nencini22}, and for finding simulation parameters predicting most realistic packing properties of membranes~\cite{antila2022emerging,NMRlipidsDatabank}. Furthermore, quantitative quality measures based on C-H bond order parameters have been recently defined and used to rank simulations in the NMRlipids databank~\cite{NMRlipidsDatabank}. However, such automatic quality evaluation is limited to simulations that have experimental data available with corresponding composition and temperature. 
%One limitation in comparing simulations of lipid mixtures with experimental data is that the various experiments are performed at different concentrations, and 
Because simulations mimicking all experimental compositions for multi-component membranes are often tedious to produce, quality evaluation of mixed lipid bilayers is not yet fully automized in the NMRlipids databank.

Here, we demonstrate how simulations of binary POPC-cholesterol mixtures can be evaluated against experimental NMR and X-ray scattering data
%order parameters and electron density profiles can be readily 
by interpolating through multiple CHOL concentrations. Because effect of cholesterol on lipid headgroup and its decoupling of acyl chain ordering has been discussed previously~\cite{botan15,antila22b}, we focus here on acyl chains that are expected to play larger role in cholesterol induced lateral membrane heterogeneity.
%to intermediate CHOL concentrations. This enables the comparison of data sets that are not extracted at the exact same CHOL concentrations, as is usually the case, especially if multiple techniques are combined in the comparison. 
We also evaluated 
%whether the simulation force fields reproduce 
the dependence of lateral diffusion coefficients on cholesterol against pulsed field gradient (PFG) NMR experiments~\cite{filippov2003effect,filippov2003influence} using the recent theoretical work that allows a quantitative comparison with experiment after eliminating the finite-size effects in MD simulations~\cite{vogele2016divergent,vogele2018hydrodynamics}. With the structural and dynamic comparisons established, we estimated the quality of popular force fields at different CHOL concentrations. We then used the best-performing force fields to analyze molecular organization at different cholesterol concentrations, as we expect them to best reflect this organization in reality.
%
%Here, we extend this work to mixtures of PC and CHOL. We compare the ability of popular force fields to reproduce effects of CHOL on the structure and dynamics of a PC lipid bilayer, and thereby 
While we focus here on POPC-cholesterol mixture, we expect our results to set guidelines for future efforts to validate intermolecular interactions in binary and more complex systems. 
%Moreover, if some of the studied force fields satisfactorily reproduce experimental behavior at least in some range of CHOL concentrations, the molecular level picture provided by it can be assumed to mimic reality in this concentration regime. 
%Thus, we compare our simulation results with different force fields to the systematic NMR data set of PC C--H bond order parameters with varying CHOL concentrations \cite{ferreira13}. In addition, we measured form factors for a range of PC--CHOL mixtures using small angle X-ray scattering, and used them to subsequently model the electron density profiles across the lipid bilayer. The form factor and electron density profiles can be compared with the same profiles calculated from the simulation data. Whereas NMR provides data on the conformations of individual lipids, form factors from scattering captures the details of the structure of the entire lipid bilayer \cite{ollila16}.

@@

\subsection{X-ray Scattering Experiments}
Fully hydrated multilamellar vesicles (MLVs), composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments, were prepared for small angle X-ray scattering (SAXS) experiments using rapid solvent exchange as described previously \cite{rieder2015optimizing,belivcka2017high} . This avoids the precipitation of cholesterol crystallites at high concentration \cite{buboltz1999novel} yielding non-phase separated samples up to 50~mol-\% cholesterol content. Lipids, purchased from Avanti Polar Lipids (Alabaster, AL, USA), were used as dry powders without any further purification. All other chemicals were obtained in pro analysis quality from Lactan (Graz, Austria). The data were obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons at $T=300$~K and analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}. The data from MLVs contain the structure factor (the crystalline lattice) and form factor in a convoluted fashion, yet by fitting the scattered intensity data we obtained both contributions. The electron density profiles were modelled from form factors using the SDP model where volume distribution functions are modelled by individual Gaussians or error functions~\cite{heberle12,Kucerka08a,kucerka12}. 
%CHOL is particularly challenging to describe within the SDP model, and therefore form factors are often compared between simulation and experiment \cite{Kucerka08b} instead of density profiles, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}. 
Cholesterol was described using two Gaussians, as proposed by Jianjun Pan (private communication), and used in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}. 
%Additional figures show the volume distribution functions and the resulting electron density profiles.
Membrane thickness was defined as twice the distance from the electron density maximum to the membrane center. The electron density maxima were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data.

@@

We performed MD simulations of a pure POPC membrane as well as five POPC/CHOL mixtures with CHOL content ranging from 11 to 47~mol-\%. Systems were simulated using four commonly used force fields, namely CHARMM36 \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig15b} models. In order to eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes (64, 256 or 1024 POPC molecules). 
%The small (64 POPC molecules) and medium (256 POPC molecules) systems were only used in the diffusion analysis, whereas all other analyses were performed on the large membranes (1024 POPC molecules). 
The number of POPC molecules was kept constant across the different CHOL concentrations. All membranes were solvated by 50 waters per lipid (POPC or CHOL).  All simulations were 1~\textmu{}s long, totaling 72~\textmu{}s, and performed using GROMACS version 2020 or 2021 \cite{pall2020heterogeneous}. The used simulation parameters are provided in Table~\ref{SItab:params}, and the simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749 (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog). All simulations were stored to the NMRlipids databank~\cite{NMRlipidsDatabank} with the ID numbers listed in Table~\ref{IDtable}.

@@

\paragraph{Structural properties:} The C--H bond order parameters, form factors and electron density profiles 
%were calculated using the \texttt{form\_factor.py} script in 
automatically calculated by the NMRlipids Databank~\cite{NMRlipidsDatabank} 
%(\url{https://github.com/NMRLipids/Databank/}) after assigning all atom types with valence electron counts. 
were used. Similarly to the experimental X-ray scattering data, membrane thickness was defined as twice the distance from the electron density maximum to the membrane center. Locations of maxima were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data. Area per phospholipid was obtained by dividing the area of the bilayer by the number of phospholipids in one leaflet. 
%
%\paragraph{Order Parameters:} The C--H order parameters were calculated as
%%
%\begin{align}
%    S_\mathrm{CD}=\left\langle\frac{3\cos^2\theta-1}{2}\right\rangle,
%\end{align}
%%
%where $\theta$ is the angle between the C--H vector and the $z$ axis normal to the bilayer plane. 
To simplify the interpolation, C-H bond order parameters of 2 (3) hydrogens in CH$_2$ (CH$_3$) groups in the POPC acyl chains were averaged. These groups rotate freely and thus the order parameters are essentially identical for both (all) hydrogens in experiments and simulations.
%, the values are also very similar, justifying this approach. 
An exception to this are the hydrogens bound to the 2\textsuperscript{nd} carbon in the oleate chain; they lack rotational averaging in both both simulations and experiments, and were thus treated separately in our analyses. The C--H order parameters for the POPC head group 
%were calculated similarly, but values for both hydrogens 
are shown separately for all CH$_2$ groups, \textit{i.e.} no averaging was performed.

@@

The diffusion coefficients extracted from the three simulation box sizes were fitted with
%
\begin{align}
	D_\mathrm{PBC}\approx D_\infty+\frac{k_\mathrm{B}T}{4\pi\mu_\mathrm{m}h}\frac{\ln\left[L/\left(L_\mathrm{SD}+1.565H\right)\right]-1.713}{1+H/L_\mathrm{SD}},
	\label{eq:sdlatpbc}
\end{align}
%
where $D_\infty$ is the lateral diffusion coefficient in an infinite system, $h$ is the hydrodynamic thickness of the membrane, $k_\mathrm{B}T$ is the Boltzmann constant, $T$ is the temperature, $H$ is half the thickness of the water layer, $L_\mathrm{SD}=\frac{h\mu_\mathrm{m}}{2\mu_\mathrm{f}}$ the Saffmann--Delbr\"{u}ck length, and $\mu_\mathrm{m}$ and $\mu_\mathrm{f}$ are shear viscosities of the membrane and the fluid (water), respectively~\cite{vogele2018hydrodynamics}. The inter-leaflet friction coefficient does not appear in Eq.~\eqref{eq:sdlatpbc} as we expect it to be infinite, which was found to be a valid assumption for lipid bilayers \cite{vogele2018hydrodynamics}. The water viscosity value of 0.3228~mPa$\cdot$s was interpolated to 298~K from the values for CHARMM TIP3P in Ref.~\citenum{ong2019temperature} and used for all simulations (CHARMM TIP3P and normal TIP3P differ by $\sim$2--3\%). The final simulation box dimensions were used for $L$ and $H$. For $h$ and $H$ ($=(L_z-h)/2$, the average values from the three systems sizes were used.

\paragraph{Quantitative Quality evaluation of cholesterol effect on membrane properties:} To ease the evaluation of simulations against experimental data with non-matching concentrations, we interpolated the effect of cholesterol in simulations and experiments for cholesterol concentrations ranging from 0\% to 50\%.
%The density profiles (as a function of distance from membrane core along the membrane normal), 
Continuous 2D matrices were created by interpolation for the oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains) and the electron density profiles (as a function of normal distance from bilayer center), while 1D interpolation was used for the cholesterol dependence of the 1\textsuperscript{st} and 2\textsuperscript{nd} form factor minima locations, and diffusion coefficients. These interpolations were then used to calculate deviations (in \%) from experimental values across CHOL concentrations to quantify the quality of the lipid force fields.  For the 2D matrices, the absolute values of the differences between matrices from simulations and experiments were first calculated. The averages of differences over the carbon atoms in the acyl chain (order parameters) or across the simulation system (density profiles) were then calculated. This resulted in 1D deviation vectors as a function of CHOL concentration. For diffusion coefficients and form factor minima, the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration. All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \% between a simulation and experiments as a function of cholesterol concentration. For C-H bond order parameters, also the deviation matrices between simulations and experiments were used to illustrate quality of simulations.

@@

@@

Cholesterol is known to induce order in cell membranes by increasing the fraction of \textit{anti} conformations in the acyl chains of phospholipids \cite{ferreira13}, which is suggested to play critical role in the phase behaviour of PC--cholesterol mixtures \cite{ipsen87}. Consequently, the correct cholesterol ordering effect is expected to be a necessary condition for a model used to understand lipid--cholesterol phase behaviour. Cholesterol ordering effect can be experimentally quantified by measuring the C-H bond order parameters using $^{13}$C or $^{2}$H NMR techniques that can be also directly compared with MD simulations~\cite{ollila16}. Simulations generally reproduce the cholesterol ordering effect, but order parameters often deviate from experiments at high cholesterol concentrations~\cite{ferreira13,madej15}. Furthermore, it has not been clear how accurately different force field parameters capture the details of lipid-cholesterol interactions and which parameters would give most realistic results for simulations of complex mixtures where such interactions play critical roles.

Here, we evaluate the cholesterol ordering effect in state of the art force field parameters against C-H bond order parameter data from $^{13}$C NMR experiments measured from POPC-cholesterol mixture with cholesterol concentratons ranging between 0-60\%~\cite{ferreira13}. To this end, we first interpolated order parameter maps as a function of acyl chain carbon number and cholesterol concentration for both simulations and experiments. These maps were then subtracted to obtain the deviation maps between simulations and experiments. The deviation maps of different force fields are shown for the palmitate (top row) and oleate (bottom row) chains of POPC in Fig.~\ref{fig:OPmaps}, whereas the original order parameter profiles are shown in Figs.~\ref{SIfig:palmitate} and \ref{SIfig:oleate}.

The cholesterol ordering effect is manifested in original profiles in Figs.~\ref{SIfig:palmitate} and \ref{SIfig:oleate} as substantial increase in absolute values of acyl chain C--H bond order parameters upon addition of cholesterol in all simulation models and experiments.
%, which can be explained by increased \textit{anti} conformations in acyl chains \cite{ferreira13,??}. 
%The absolute values of order parameters of \textit{sn}-1 chain from $^{13}$C NMR experiments \cite{ferreira13} exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig. \ref{slopes}), which is expected because phase separation is not observed for this mixture \cite{ionova12,ferreira13}.
The deviations mapped in Fig.~\ref{fig:OPmaps} provide an intuitive view for a quantitative comparison of different force fields with experiments. In white regions, the simulation results are considered to fall within the experimental error as the deviations are in the range of [--0.02,0.02]. Blue indicates that the order parameters are too negative, i.e., the acyl chains are too ordered in simulations, and \textit{vice versa} for red. Overall, the simulation models behave reasonably well at low CHOL concentrations, but deviate significantly from experiment at higher CHOL concentrations.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\linewidth]{../FIGS/OP_chains.pdf}
  \caption{\label{fig:OPmaps}%
  POPC acyl chain order parameter deviation from experiments. Data are shown for palmitate (top row) and oleate (bottom row) and for the four force fields (columns). Negative values indicate that order is too high ($-S_\mathrm{CH}$ values too negative) in the simulations. The values that are within the estimated experimental error range of 0.02~\cite{ollila16} are 
  %considered to be ``correct'' and 
  coloured in white. Order parameters of identical hydrogens attached to the same carbons were averaged with the exception of C2 carbon of the oleate chain with the forked order parameters for which differences for both the larger and smaller values were calculated, and the average of this deviation is shown in the C2 column of oleate.
  }
\end{figure}

In CHARMM36, the both the palmitate and oleate chains get too ordered upon increasing CHOL concentration, yet the oleate C9 carbon next to the double bond is too disordered in the simulation. The oleate chain shows best agreement with experiment in Slipids with only slight excess ordering at higher CHOL concentrations. The excess ordering is more severe for the palmitate chain. Still, the major discrepancy between Slipids and experiment is the drastically too disordered C2 and C3 carbons of the palmitate chain. This effect was not present in the original Slipids model \cite{NMRlipidsDatabank}, indicating that it was introduced in the recent reparametrization that improved the head group and glycerol backbone structures of Slipids \cite{grote2020optimization}. Lipid17 provides the best overall agreement with experiment, as no segments deviate significantly from experiment at any CHOL concentrations. At low CHOL concentrations, the middle of the oleate chain is too disordered.
%, which agrees with the large APL observed in Fig.~\ref{fig:densmaps}. 
At large CHOL concentrations, the oleate chain agrees well with experiment, whereas the palmitate chain has a slightly too ordered C9--C12 segment, yet the deviation is significantly smaller than in the other force fields. MacRog behaves reasonably well at low CHOL concentrations, yet the C2--C5 carbons in palmitate are slightly too disordered, whereas the C2 of oleate is somewhat too ordered. However, at larger CHOL concentrations the chains become overall way too ordered, leading to the largest overall deviations from experiment.

%The head group order parameters are not largely affected by CHOL concentration in experiments \cite{ferreira13}, which has been discussed in our earlier work \cite{botan15}. The four force fields evaluated here also reproduce this behavior to a great degree (Fig.~\ref{SIfig:headgroups}), so the limitations observed in the absence of CHOL \cite{botan15} largely persist across all studied CHOL concentrations. Moreover, it has now become standard to consider them in the parameterization and validation of lipid force fields \cite{dickson2022lipid21,grote2020optimization,yu2021charmm36,yu2021semi}.

\subsection{Cholesterol Effect on membrane properties is Manifested Differently in Different Force Fields}
